Untitled

### 2015-03-26 leaflifespan.Rmd
work on scripts

#discuss with Daniel
- once we are certain about leaf lifespan calculations, those calculations should go into a separate file, since that is background processing, but like calculating investment. Then this becomes the file where variables are summarized, plotted analyzed.
- make sure regressions exist (or are done), to separately look at leaf and stem biomass
- seed weight vs. propagule weight dilemma
- sites that are aged 1.35 vs 1.4 years

@@

@@

# load packages and "open" dataframes to use
library(remake)
create_bindings()

LMA <- filter(LMA_raw, species!="" & species!=" ")

wood <- filter(woodDensity_raw, use=="use")

seedsize <- seedSize_raw

harvest <- filter(HarvestData, use_status=="use" & segment==1 & start_end=="end")

investment <- ReproductionAllocation_all
names(investment) <- c("individual","ReproInv","age","GrowthInv","StartWeight","FinalWeight","StartBasalDiamAv","FinalBasalDiamAv","TotalInvestment","RA","species")
#still have tagId as label
#remaking this from scratch

accessory <- AccessoryCosts_all

leafLifespan <- filter(leafLifespan_raw, dont_use!="dead" & dont_use!="dont_use")
leafLifespan$individual <- as.factor(leafLifespan$individual)
#loading this column as character, so using this for now

@@

leaf_lifespan_temp <- leafLifespan %>%
              select(species,age,site,individual,d_start,d_end,length_start,new_length,segment,lvs_start_count,lvs_end_count,lvs_new_count,new_but_shed_count)

@@

#adding investment and accessory costs data to leafLifespan dataframe to create a dataframe with all individual level data
#SummaryInd <- left_join(leafLifespan, investment, by = "individual")
#SummaryInd <- left_join(SummaryInd, accessory, by = "individual")
#SummaryInd <- left_join(SummaryInd, harvest, by = "individual")
#not sure why the above text isn't working... using merge instead

SummaryInd <- merge(leafLifespan, investment, by.x=c("individual"), by.y=c("individual"))
SummaryInd <- subset(SummaryInd, select = -c(age.y,species.y) )
SummaryInd <- merge(SummaryInd, accessory, by.x=c("individual"), by.y=c("individual"))
SummaryInd <- subset(SummaryInd, select = -c(age,species) )
SummaryInd <- merge(SummaryInd, harvest, by.x=c("individual"), by.y=c("individual"))
#problem here, because duplicates number of entries
SummaryInd <- subset(SummaryInd, select = -c(age.x,species.x) )

@@

#TODO: Check outlier for HATE

@@

#harvest summary by species, age
harvest_summary <- harvest %>%
  group_by(species, age2) %>%
  summarise_each(funs(mean, se, length), total_stem_weight, total_leaf_weight, height)

@@

#wood density summary by species
Wood_summary <- wood %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), density)
names(Wood_summary)<-c("species","WD_mean","WD_se","WD_length")

@@

#maximum height for each species
MaxH_spp <- harvest %>%
  group_by(species) %>%
  summarise_each(funs(max), height)
names(LMA_spp)<-c("species", "maxH_spp")

SummaryInd <- merge(SummaryInd,Wood_summary, by.x=c("species"), by.y=c("species"))
SummaryInd <- merge(SummaryInd,LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
SummaryInd <- merge(SummaryInd,MaxH_spp, by.x=c("species"), by.y=c("species"))
SummaryInd <- merge(SummaryInd,seedsize, by.x=c("species"), by.y=c("species"))
SummaryInd$leafArea <- SummaryInd$total_leaf_weight * SummaryInd$LMA_mean
#TO DO: REMOVE DUPLICATE COLUMNS FOR AGE and SPECIES MORE ELEGANTLY
#TO DO: REMOVE COLUMNS I DON'T NEED

#merge various data summaries into SummarySppAge dataframe
SummarySppAge <- merge(LL_summary, LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, Wood_summary, by.x=c("species"), by.y=c("species"))
SummarySppAge <- merge(SummarySppAge, seedsize, by.x=c("species"), by.y=c("species"))
SummarySppAge <- merge(SummarySppAge, MaxH_spp, by.x=c("species"), by.y=c("species"))
SummarySppAge <- merge(SummarySppAge, maxH, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, investment_summary, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, accessory_summary, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, harvest_summary, by.x=c("species","age"), by.y=c("species","age"))
#TO DO: REMOVE DUPLICATE COLUMNS FOR AGE

#merge various data summaries into SummarySpp dataframe
SummarySpp <- merge(LL_spp, LMA_spp, by.x=c("species"), by.y=c("species"))
SummarySpp <- merge(SummarySpp, Wood_summary, by.x=c("species"), by.y=c("species"))
SummarySpp <- merge(SummarySpp, seedsize, by.x=c("species"), by.y=c("species"))
SummarySpp <- merge(SummarySpp, maxH, by.x=c("species"), by.y=c("species"))

#WONDERING IF I SHOULD AT THIS POINT HAVE THE DATA FRAME CALLED "sUMMARYINDIVIDUAL" INSTEAD, SINCE NOW CONTAINS FAR MORE THAN LL DATA

@@

```{r, echo=FALSE,results='hide', message=FALSE}
mod <- lm(LL_birth ~ age + species +age*species, data=subset(SummaryInd,age>3 & LL_birth!="Inf"))
summary(mod)

@@

@@

@@

#models looking at how the relationship between leaf weight (or area) and total plant weight shifts with age
this should be another way to get at the decreasing "leafiness" of plants with age

```{r, echo=FALSE,results='hide', message=FALSE}
plot(leaf_weight~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(FinalWeight~leaf_weight+age+species,SummaryInd)
summary(mod)

plot_LeafWeight <- function(spp) {
  plot(leaf_weight~FinalWeight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(FinalWeight~leaf_weight+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

@@

>>>>

#discuss with Daniel
- once we are certain about leaf lifespan calculations, those calculations should go into a separate file, since that is background processing, but like calculating investment. Then this becomes the file where variables are summarized, plotted analyzed.
- add final leaf and stem biomass output to "investment" output file
- seed weight vs. propagule weight dilemma

@@

#data to check
- HATE outlier in LL

@@

# load packages and "open" dataframes to use
library(remake)
create_bindings()
```

```{r, echo=FALSE,results='hide', message=FALSE}
LMA <- filter(LMA_raw, species!="" & species!=" ") %>%
  select(species,age,LMA)

wood <- filter(woodDensity_raw, use=="use") %>%
  select(species,density)

seedsize <- seedSize_raw %>%
  select(species,seed_size)

harvest <- filter(HarvestData, use_status=="use" & segment==1 & start_end=="end") %>%
  select(species,individual,age,age2,height,total_leaf_weight,site)

investment <- ReproductionAllocation_all
investment$age2 <- investment$age
investment$age <- round(investment$age, digits=1)

accessory <- AccessoryCosts_all
accessory$age2 <- accessory$age
accessory$age <- round(accessory$age, digits=1)
#accessory$PerAccessory <- {
#  select(accessory, Total >0,
#  1 - (accessory$Prop/accessory$Total)
#}

#accessory$PerPropagule = 1 - accessory$PerAccessory

leafLifespan <- filter(leafLifespan_raw, dont_use!="dead" & dont_use!="dont_use") %>%
  select(species,age,individual,d_start,d_end,length_start,new_length,lvs_start_length,lvs_start_count,lvs_end_length,lvs_end_count,lvs_new_length,lvs_new_count,new_but_shed_count,mm_lvs_spec,count_lvs_spec)

leafLifespan$individual <- as.factor(leafLifespan$individual)
leafLifespan$species <- as.factor(leafLifespan$species)

#loading this column as character, so using this for now

@@

leaf_lifespan_temp <- leafLifespan %>%
              select(species,age,individual,d_start,d_end,length_start,new_length,lvs_start_count,lvs_end_count,lvs_new_count,new_but_shed_count)

@@

#adding investment and accessory costs data to leafLifespan dataframe to create a dataframe with all individual level data
SummaryInd <- merge(leafLifespan, select(investment, -species, -age), by="individual")
SummaryInd <- merge(SummaryInd, select(accessory, -species, -age, -age2), by="individual")
SummaryInd <- merge(SummaryInd, select(harvest, -species, -age, -age2), by="individual")

@@

@@

#harvest summary by species, age
harvest_summary <- harvest %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), height,total_leaf_weight)
#names(harvest_summary)<-c("species","age","height_mean","height_se","height_length")

@@

#wood density summary by species
wood_summary <- wood %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), density)
names(wood_summary)<-c("species","WD_mean","WD_se","WD_length")

@@

#maximum height for each species
maxH_spp <- harvest %>%
  group_by(species) %>%
  summarise_each(funs(max), height)
names(maxH_spp)<-c("species", "maxH_spp")

SummaryInd <- merge(SummaryInd,wood_summary, by=c("species"))
SummaryInd <- merge(SummaryInd,LMA_summary, by=c("species","age"))
SummaryInd <- merge(SummaryInd,maxH_spp, by=c("species"))
SummaryInd <- merge(SummaryInd,seedsize, by=c("species"))

SummaryInd$leafArea <- SummaryInd$total_leaf_weight * SummaryInd$LMA_mean

#merge various data summaries into SummarySppAge dataframe
SummarySppAge <- merge(LL_summary, LMA_summary, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, wood_summary, by=c("species"))
SummarySppAge <- merge(SummarySppAge, seedsize, by=c("species"))
SummarySppAge <- merge(SummarySppAge, maxH_spp, by=c("species"))
SummarySppAge <- merge(SummarySppAge, maxH, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, investment_summary, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, accessory_summary, by=c("species","age"))
SummarySppAge <- merge(SummarySppAge, harvest_summary, by=c("species","age"))

#merge various data summaries into SummarySpp dataframe
SummarySpp <- merge(LL_spp, LMA_spp, by=c("species"))
SummarySpp <- merge(SummarySpp, wood_summary, by=c("species"))
SummarySpp <- merge(SummarySpp, seedsize, by=c("species"))
SummarySpp <- merge(SummarySpp, select(maxH, -age), by=c("species"))

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(LL_birth~age,SummaryInd,log="xy", xlab="age",ylab="leaf lifespan (based on leaf births",col=col.spp[as.factor(species)])
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)

mod <- lm(LL_birth ~ age + species, data=subset(SummaryInd,age>3 & LL_birth!="Inf"))
summary(mod)

@@

plot(LMA_mean~age,SummarySppAge,log="xy", xlab="age",ylab="LMA",col=col.spp[as.factor(species)],pch=16)
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(lvs_end_total~age,SummaryInd,log="xy", xlab="age",ylab="final leaf count along shoot",col=col.spp[as.factor(species)])
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
mod <- lm(lvs_end_total~age+species+age*species,SummaryInd)
summary(mod)

@@

#other models looking at how the relationship between leaf weight (or area) and total plant weight shifts with age
this should be another way to get at the decreasing "leafiness" of plants with age

```{r, echo=FALSE,results='hide', message=FALSE}
plot(total_leaf_weight~FinalWeight,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=0.8)
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <-lm(FinalWeight~total_leaf_weight+age+species,SummaryInd)
summary(mod)

plot_LeafWeight <- function(spp) {
  plot(total_leaf_weight~FinalWeight,data=subset(SummaryInd,species==spp), xlab="total weight (mg)", ylab ="leaf weight (mg)", log="xy",col=col.age[as.factor(age)], pch=16,     cex=0.8,main=labels.spp[spp])
  legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
  mod <-lm(FinalWeight~total_leaf_weight+age,data=subset(SummaryInd,species==spp))
  abline(mod)
  summary(mod)
}

@@

#Plots with RA by age 
Will include a couple of the obviously different species in lab meeting

```{r, echo=FALSE,results='hide', message=FALSE}
plot_RAbyAge <- function(spp) {
  plot(RA~age, data=subset(SummaryInd,species==spp), xlab="age",ylab="reproductive allocation", log="x", main=labels.spp[spp])
  points(RA_mean~age, data=subset(SummarySppAge, species==spp),pch=16,cex=1.2)
}

for(n in names(labels.spp)) {
  plot_RAbyAge(n)
} 
```

#Plots of Accessory costs by age
```{r, echo=FALSE,results='hide', message=FALSE}

plot_PropagulebyAge <- function(spp) {
  plot(Prop~age, data=subset(SummaryInd,species==spp), xlab="age",ylab="investment on propagule", log="x", main=labels.spp[spp])
  points(Prop_mean~age, data=subset(SummarySppAge, species==spp),pch=16,cex=1.2)
}

for(n in names(labels.spp)) {
  plot_PropagulebyAge(n)
}
